First Season MWA EoR Power Spectrum Results at Redshift 7

### 2016-08-02 ms.tex
nearly finished with minor referee comments

After calibration, we form snapshot image cubes (frequency mapping to line-of-sight 
direction). The visibilities are gridded using the primary beam as the gridding kernel, placing 
the data in the holographic frame, which has been shown to be the optimal frame for 
combining images in the sense that it preserves all information for parameter estimation 
\citep{Morales:2009,Bhatnagar:2008}. We use a simulation-based primary beam model for MWA 
antennas developed by \citet{Sutinjo:2015} 
which incorporates an average embedded element pattern for the dipoles, as well as mutual
coupling between dipoles within an antenna.
At this stage we also average in frequency by 
a factor of two by gridding pairs of frequency channels to the same $(u,v)$ plane, resulting in 
a frequency resolution of 160 kHz.

@@

After squaring and dividing by our observation window function \citep{Bowman:2006}, we 
arrive in three dimensional power spectrum space. We use the even/odd interleaved cubes 
to form a signal (odd plus even) and noise (odd minus even) power spectrum, which we 
subtract from one another to form an unbiased estimate of our signal power. This is 
mathematically equivalent to cross multiplying the even and odd cubes. The three 
dimensional power spectrum cube can next be averaged in annuli to form two dimensional 
power spectra, or spherical shells to form one dimensional power spectra where we will 
ultimately constrain the EoR.

\begin{figure}
\begin{center}
\includegraphics[width=\columnwidth]{example_ps.pdf}
\caption{
An example two dimensional, E-W polarization residual power spectrum formed using the 
three hour golden data set
from August 23, 2013. The foreground wedge dominates the region below the solid 
black horizon line, while regions above are mostly noise-like with the exception of horizontal 
coarse band harmonic lines. 
The vertical streaks at high $k_{||}$ and high $k_{\perp}$ are due to sparse $(u,v)$ sampling
beyond the dense core of the MWA.
\label{fig:example_ps}
}
\end{center}
\end{figure}

An example two dimensional residual power spectrum formed from the three hour golden 
data set is shown in Figure~\ref{fig:example_ps}. The bulk of the residual power is in the 
so-called foreground ``wedge", indicated with the diagonal black lines where the solid line 
corresponds to sky emission at the horizon, and the dashed lines corresponds to emission 
at the edge of the MWA field of view. Above the wedge we see horizontal lines of 
contamination due to the periodic frequency sampling function. These lines are often 
referred to as the coarse band harmonics. We see vertical streaks at high $k_{||}$ and high 
$k_{\perp}$. These are due to sparse $(u,v)$ sampling beyond the dense core of the MWA 
(starting $\sim70\lambda$), and the foreground power mixing to high spectral 
modes \citep{Bowman:2009}. Between the coarse band lines and to the left of the vertical streaks we see regions 
which contain both positive and negative (indicated by blue on the color bar) pixels. These 
regions are where we have successfully isolated the foregrounds and retained a noise-like 
EoR window in the three hour integration.

@@

Our pipeline uses two modes of FHD -- the full deconvolution mode to identify point sources 
and build a catalog, and a production mode where we calibrate using the sky model and 
subtract it from the data without further fitting. Because full deconvolution on every 
observation from the MWA is computationally not feasible, we currently restrict ourselves to 
the golden data set to build our model. The process of building the catalog is presented in 
\citet{Carroll:2016}. They applied machine learning classification methods 
to select reliable detections from the full deconvolution FHD mode, culminating in the KGS 
catalog. The work here uses an early iteration of the KGS catalog that was available at the 
time of analysis.

@@

Because we use the input foreground model to calibrate our visibilities, the gain estimates 
for our two runs differed. In particular we saw the overall flux scale of the calibrated 
visibilities was higher when using the full hierarchical catalog. In order to put both residual 
2D power spectra onto the same scale for comparison, we first divide each pixel by the 
corresponding dirty power spectrum pixel, then subtract to arrive at a ratio difference. 
\begin{equation}\label{eq:diff_ratio}
\mathtt{ratio\;difference = \frac{residual_1}{dirty_1} - \frac{residual_2}{dirty_2}}
\end{equation}
Here we use subscript 1 to represent the MWACS catalog spectra, and 2 for the 
hierarchical catalog spectra. The ratio difference is shown in Figure~\ref{fig:mwacs_ratio}. 
The entire wedge being positive (blue) demonstrates that using the new catalog resulted in 
successfully subtracting higher fractional power. Not only does this indicate we subtract 
more power, but it also confirms that our calibrated data is more closely matched to the 
model, meaning that our calibration solutions in general are more accurate.

@@

This first iteration of modeling the diffuse structure within our primary field was done by 
simply using output point-source subtracted residual images of FHD from the three hour 
golden data set. We combine the images from three hours to leverage the rotation of the 
earth to improve the point spread function (PSF) of the instrument. This integrating of 
images is done in a uniformly weighted frame to minimize the effect of double instrument 
convolution (once when the data are taken, again when we use the output as a model). We 
then form a pseudo Stokes I image by adding beam-weighted East-West and North-South 
polarizations. We also integrate in frequency to form a single continuum image for our 
model. By forcing our model to contain no spectral structure we mitigate the risk of 
subtracting cosmological signal.

@@

We demonstrate the impact of our diffuse model again by running the golden data set 
through our entire analysis pipeline with and without using the diffuse model. Because the 
foreground models used for calibration in the two runs are identical (diffuse is omitted for 
calibration), the calibration solutions were identical and therefore the dirty power spectra 
were identical as well. We compare the model and residual power spectra by direct 
subtraction in Figure~\ref{fig:diffuse_diff}. The top panel shows the model power with the 
diffuse subtracted from the model power without. 
Because our diffuse model added power to the full foreground model, the entire plot is
negative (red).
Even the EoR window is red (albeit at a much lower level than the wedge) because of power
leakage from non-uniform spectral sampling and from the dynamic range limit of the 
Blackman-Harris window function.
The bottom panel is the difference in residual 
power. The wedge is completely positive (blue), indicating that the diffuse model 
successfully subtracted from the dirty visibilities. The differences in the window are positive 
and negative, indicating that 
the differences in the foreground model in this region are below the three hour noise level.

\begin{figure}
\begin{center}
\subfigure{\includegraphics[width=\columnwidth]{diffuse_diff_model.pdf}}
~
\subfigure{\includegraphics[width=\columnwidth]{diffuse_diff_residual.pdf}}
\caption{
Single polarization (E-W) power spectrum differences from the golden data set showing the 
effect of subtracting a diffuse foreground model. The calibration is identical with and without 
the diffuse model, so the dirty power spectra are identical. The diffuse model adds a large 
amount of power to the model (top), and the residual power (bottom) is significantly lower in 
the wedge when using the diffuse model (positive difference), demonstrating a successful 
subtraction. While this first attempt at a diffuse model was rudimentary, it shows promise as 
it successfully removed 90\% of the residual power. 
\label{fig:diffuse_diff}
}
\end{center}
\end{figure}

>>>>

After calibration, we form snapshot image cubes (frequency mapping to line-of-sight 
direction). The visibilities are gridded using the primary beam as the gridding kernel, placing 
the data in the \textbf{``holographic frame,"} which has been shown to be the optimal \textbf{weighting scheme} for 
combining images in the sense that it preserves all information for parameter estimation 
\citep{Morales:2009,Bhatnagar:2008}. We use a simulation-based primary beam model for MWA 
antennas developed by \citet{Sutinjo:2015} 
which incorporates an average embedded element pattern for the dipoles, as well as mutual
coupling between dipoles within an antenna.
At this stage we also average in frequency by 
a factor of two by gridding pairs of frequency channels to the same $(u,v)$ plane, resulting in 
a frequency resolution of 160 kHz.

@@

\textbf{The pipeline as described up to this point is applied to both the calibrated data
and the foreground model. At this point we subtract the model from the data to form
residual data as well. All three sets of data (dirty, model, and residual) are carried through 
the remaining steps to form corresponding power spectra.}

After squaring and dividing by our observation window function \citep{Bowman:2006}, we 
arrive in three dimensional power spectrum space. We use the even/odd interleaved cubes 
to form a signal (odd plus even) and noise (odd minus even) power spectrum, which we 
subtract from one another to form an unbiased estimate of our signal power. This is 
mathematically equivalent to cross multiplying the even and odd cubes. The three 
dimensional power spectrum cube can next be averaged in annuli 
\textbf{of constant $k_{\perp}$ and $k_{||}$ (i.e. orthogonal to the $k_{||}$ axis)}
to form two dimensional 
power spectra, or spherical shells to form one dimensional power spectra where we will 
ultimately constrain the EoR.

\begin{figure}
\begin{center}
\includegraphics[width=\columnwidth]{example_ps.pdf}
\caption{
An example two dimensional, E-W polarization residual power spectrum formed using the 
three hour golden data set
from August 23, 2013. The foreground wedge dominates the region below the solid 
black horizon line, while regions above are mostly noise-like with the exception of horizontal 
coarse band harmonic lines. 
The vertical streaks at high $k_{||}$ and high $k_{\perp}$ are due to sparse $(u,v)$ sampling
beyond the dense core of the MWA.
\textbf{The right axis shows the delay (Fourier dual to frequency) corresponding to the 
$k_||$ axis, and the top axis shows the baseline length (in wavelengths) corresponding
to the $k_{\perp}$ axis.}
\label{fig:example_ps}
}
\end{center}
\end{figure}

An example two dimensional residual power spectrum formed from the three hour golden 
data set is shown in Figure~\ref{fig:example_ps}. The bulk of the residual power is in the 
so-called foreground ``wedge", indicated with the diagonal black lines where the solid line 
corresponds to sky emission at the horizon, and the dashed lines corresponds to emission 
at the edge of the MWA field of view. Above the wedge we see horizontal lines of 
contamination due to the periodic frequency sampling function. These lines are often 
referred to as the coarse band harmonics. We see vertical streaks at high $k_{||}$ and high 
$k_{\perp}$. These are due to sparse $(u,v)$ sampling beyond the dense core of the MWA 
(starting $\sim70\lambda$), and the foreground power mixing to high spectral 
modes \citep{Bowman:2009}. 
\textbf{We provide axes more closely related to the measurements to assist in connecting
these instrumental effects. The right axis shows delay (Fourier dual to frequency), and the
top axis shows baseline length measured in wavelengths.}
Between the coarse band lines and to the left of the vertical streaks we see regions 
which contain both positive and negative (indicated by blue on the color bar) pixels. These 
regions are where we have successfully isolated the foregrounds and retained a noise-like 
EoR window in the three hour integration.

@@

Our pipeline uses two modes of FHD -- the full deconvolution mode to identify point sources 
and build a catalog, and a production mode where we calibrate using the sky model and 
subtract it from the data without further fitting. Because full deconvolution on every 
observation from the MWA is computationally not feasible, we currently restrict ourselves to 
the golden data set to build our model. The process of building the catalog is presented in 
\citet{Carroll:2016}. They applied machine learning classification methods 
to select reliable detections from the full deconvolution FHD mode, culminating in the KGS 
catalog\footnote{\textbf{KGS is an abbreviation of KATALOGSS, the KDD (Knowledge 
Discovery in Databases) Astrometry, Trueness, and Apparent Luminosity of Galaxies in 
Snapshot Surveys}}. The work here uses an early iteration of the KGS catalog that was available at the 
time of analysis.

@@

Because we use the input foreground model to calibrate our visibilities, the gain estimates 
for our two runs differed. In particular we saw the overall flux scale of the calibrated 
visibilities was higher when using the full hierarchical catalog. In order to put both residual 
2D power spectra onto the same scale for comparison, we first divide each 
\textbf{$(k_{\perp},k_{||})$ cell}
by the 
corresponding dirty power spectrum pixel, then subtract to arrive at a ratio difference. 
\begin{equation}\label{eq:diff_ratio}
\mathtt{ratio\;difference = \frac{residual_1}{dirty_1} - \frac{residual_2}{dirty_2}}
\end{equation}
Here we use subscript 1 to represent the MWACS catalog spectra, and 2 for the 
hierarchical catalog spectra. The ratio difference is shown in Figure~\ref{fig:mwacs_ratio}. 
The entire wedge being positive (blue) demonstrates that using the new catalog resulted in 
successfully subtracting higher fractional power. Not only does this indicate we subtract 
more power, but it also confirms that our calibrated data is more closely matched to the 
model, meaning that our calibration solutions in general are more accurate.

@@

This first iteration of modeling the diffuse structure within our primary field was done by 
simply using output point-source subtracted residual images of FHD from the three hour 
golden data set. We combine the images from three hours to leverage the rotation of the 
earth to improve the point spread function (PSF) of the instrument. This integrating of 
images is done \textbf{with uniform weighted data} to minimize the effect of double instrument 
convolution (once when the data are taken, again when we use the output as a model). We 
then form a pseudo Stokes I image by adding beam-weighted East-West and North-South 
polarizations. We also integrate in frequency to form a single continuum image for our 
model. By forcing our model to contain no spectral structure we mitigate the risk of 
subtracting cosmological signal.

@@

We demonstrate the impact of our diffuse model again by running the golden data set 
through our entire analysis pipeline with and without using the diffuse model. Because the 
foreground models used for calibration in the two runs are identical (diffuse is omitted for 
calibration), the calibration solutions were identical and therefore the dirty power spectra 
were identical as well. We compare the model and residual power spectra by direct 
subtraction in Figure~\ref{fig:diffuse_diff}. 
The top panel shows \textbf{our foreground model power spectrum
minus the model power spectrum when including diffuse}. 
Because our diffuse model added power to the full foreground model, the entire plot is
negative (red).
Even the EoR window is red (albeit at a much lower level than the wedge) because of power
leakage from non-uniform spectral sampling and from the dynamic range limit of the 
Blackman-Harris window function.
The bottom panel is the difference in residual 
power. The wedge is completely positive (blue), indicating that the diffuse model 
successfully subtracted from the dirty visibilities. The differences in the window are positive 
and negative, indicating that 
the differences in the foreground model in this region are below the three hour noise level.

\begin{figure}
\begin{center}
\subfigure{\includegraphics[width=\columnwidth]{diffuse_diff_model.pdf}}
~
\subfigure{\includegraphics[width=\columnwidth]{diffuse_diff_residual.pdf}}
\caption{
Single polarization (E-W) power spectrum differences from the golden data set showing the 
effect of subtracting a diffuse foreground model. The calibration is identical with and without 
the diffuse model, so the dirty power spectra are identical. 
\textbf{In the top panel the model power spectrum difference is almost entirely negative 
(red), indicating that the diffuse model added a large amount of power to the foreground 
model. The bottom panel shows the residual power spectrum difference is positive in the 
wedge (blue), demonstrating a successful subtraction.}
While this first attempt at a diffuse model was rudimentary, it 
\textbf{successfully removed 90\% of the residual power. }
\label{fig:diffuse_diff}
}
\end{center}
\end{figure}
